Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-31 src/ms.tex
revising Gaia appendix

\citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies. We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample. Such an offset could be due to differing spectroscopic temperature scales between the two studies. To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019}. We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average. Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.

@@

\edit1{We note that the morphology of the color–period distribution appears to be different between the Gaia–McQuillan and Gaia–Santos catalogs, with the Gaia–Santos distribution presenting a break near $G_{BP}-G_{RP}= 0.7$. This break is not apparent in the Gaia–McQuillan sample. Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}. At present, we do not have a satisfactory explanation for this behavior.}

>>>>

\citet{Hall2021} determined rotation periods for 91 main-sequence, asteroseismic Kepler targets from rotational splitting of asteroseismic oscillation frequencies. We found that the distribution of the asteroseismic sample in the \teff--\prot plane approximately matches the pile-up we observe, although the \citet{Hall2021} sample appears shifted slightly to either higher \prot or higher \teff values relative to the ridge in the \edit1{\lamostmcq} sample while such an offset is either absent or not as apparent relative to the CKS sample. Such an offset could be due to differing spectroscopic temperature scales between the two studies, \edit1{differing observational biases, or some combination of the two effects}.

To investigate the origin of this offset, we cross-matched the \citet{Hall2021} sample with the LAMOST DR5 catalog \citep{Xiang2019} \edit1{and the CKS sample}. We found a root mean square (RMS) of the residuals between the \citet{Hall2021} and LAMOST DR5 \teff measurements of 55~K, with a median offset of 41~K, such that the LAMOST \teff scale is cooler, on average (see Appendix~\ref{app:teff}). \edit1{Similarly, we found excellent agreement between the \teff scales of the CKS and \citet{Hall2021} samples, with a median offset of 29~K (such that the CKS scale is hotter) and an RMS of 41~K.}

{\color{red} Though this discrepancy is modest, adopting the LAMOST \teff scale appeared to resolve most of the offset between the pile-up as seen in the two samples.}

@@

\edit1{We note that the morphology of the color–period distribution appears to be different between the Gaia–McQuillan and Gaia–Santos catalogs, with the Gaia–Santos distribution presenting a break near $G_{BP}-G_{RP}= 0.7$. This break is not apparent in the Gaia–McQuillan sample. Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}, suggesting the origin of the discontinuity is in the rotation periods rather than the temperatures or colors. At present, we do not have a satisfactory explanation for this behavior, though we note that the \citet{Santos2021} catalog employed various time series filters which might introduce systematic artifacts.}
